---
title: "R Notebook"
output: html_notebook
---

## The objectives of the lab

The purpose of this lab is to reproduce tables from the third chapter of the book "Elements of
Statistical Learning" from Hastie, Tibshirani and Friedman, as they are shown below. (Table 3.1 and 3.2)

###Ex 1 - Tables 3.1 and 3.2

1) Prepare the data  
  a) Raw data is available on the web at http://statweb.stanford.edu/~tibs/ElemStatLearn.1stEd/datasets/prostate.data . It has already been downloaded and is in the *data* folder.
  
```{r}
data <- read.table("data/prostate.data", sep = "")
head(data)
```
b) Extract and normalize the explicative variables

```{r}
X <- scale(data[,1:8])
```
c) Is it wise to normalize these data?
```{r}
summary(data)
```
In ordinary linear regression, we sometimes normalize data for numerical reasons, when they differ by a large order of magnitude. Here the ranges of the different variables are quite similar so it is not necessary to normalize. However here normalization was done by the textbook. Since we want to reproduce the tables of the textbook, we normalize our data.

```{r}
X <- scale(data[,1:8])
```

d) Extract the target variable

```{r}
Y <- as.matrix(data[,"lpsa"])
```

e) Split the data into train and test set

```{r}
Xtrain <- X[data[["train"]], ]
Ytrain <- Y[data[["train"]],]
Xtest <- X[!data[["train"]], ]
Ytest <- Y[!data[["train"]], ]
```

2. Compute the correlations of predictors in the prostate cancer data as presented in  **Table 3.1**

```{r}
Xtrainscale <- scale(Xtrain)
C <- cov(as.matrix(Xtrainscale))
C
```

3. Reproduce the results presented Table 3.2
  a) Compute the coefficients of the linear regression model, without using the lm function
(but you can use it validate your code)

```{r}
Xtrainone <- cbind(array(1, dim = c(nrow(Xtrain),1)), Xtrain)
b <- solve(t(Xtrainone) %*% Xtrainone, t(Xtrainone) %*% Ytrain)
```

  b) Compute the prediction error
  
```{r}
Ypred <- Xtrainone %*% b
err <- Ytrain - Ypred
```

  c) Compute the standard error for each variable
  
```{r}
sig2 <- (t(err) %*% err)/ (nrow(Xtrainone) - ncol(X) -1)
v <- diag(solve(t(Xtrainone) %*% Xtrainone))
stderr <- sqrt(as.vector(sig2)) * sqrt(v)
```
  
  d) compute the Z score for each variable
```{r}
Z <- b/stderr
```

  e) visualize the results and compare with table 3.2

```{r}
table32 <- cbind(b,stderr,Z)
table32
```

**Ex. 2 — Your turn**

Reproduce Table 3.3, at least the first four columns that is LS, Best Subset, Ridge and Lasso.

We already have the LS column, this corresponds to our **b** vector computed earlier.

For the *Best subset* column, we need to find the best model among all simple linear regression models (we do not consider models with interactions)
For now to make things easier, let's assume we know the best model only contains the lcavol and the lweight coefficients.

```{r}
Xb <- X[,c("lcavol", "lweight")]
```

Split into test and train sets
```{r}
Xb_train <- Xb[data[["train"]], ]
Xb_test <- Xb[!data[["train"]], ]

#Uncomment the next line if you need to scale again
#Xbest_trainscale <- scale(Xbest_train)
```

```{r}
#Add 1s to our feature matrix
Xb_trainone <- cbind(array(1, dim = c(nrow(Xb_train),1)), Xb_train)
#Solve normal eqns and find b_best
b_best <- solve(t(Xb_trainone) %*% Xb_trainone, t(Xb_trainone) %*% Ytrain)

#Compute prediction error on the train set for each data point
Yb_pred <- Xb_trainone %*% b_best
errb <- Ytrain - Yb_pred

#Compute standard error for each variable
sig2b <- (t(errb) %*% errb)/ (nrow(Xb_trainone) - ncol(X) -1)
vb <- diag(solve(t(Xb_trainone) %*% Xb_trainone))
stderrb <- sqrt(as.vector(sig2)) * sqrt(v)

#Compute prediction error on the test set for each data point
Xb_testone <- cbind(array(1, dim = c(nrow(Xb_test),1)), Xb_test)
Yb_pred_test <- Xb_testone %*% b_best
errb_test <- Ytest - Yb_pred_test
```
Create the column *Best Subset*
```{r}
best_subset <- c(b_best, matrix(0, 6, 1), mean(errb_test), sd(errb_test))
```

We could also fit all the models to find the best subset.

```{r}
dummies = expand.grid(0:1, 0:1, 0:1, 0:1, 0:1, 0:1, 0:1, 0:1)
dim(dummies)
```


Let's compute the Ridge column.

```{r}

```

